Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-05-31 mu/article/article.tex
Completed plots for Cal V_{s^z}
V_{s^z} layer description completed

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
                        % documents and URLs 
\usepackage{blindtext}  % fill text
\usepackage{color}
\usepackage{multirow}
% \usepackage{showframe}
\usepackage[inline]{showlabels}

@@

\section{Introduction}
\blindtext
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/altstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/altstruc1}
    \caption{Alt structure.}
    \label{fig:alt-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/upstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/upstruc1}
    \caption{Up structure}
    \label{fig:up-struc}
\end{figure}
\blindtext
\blindtext

@@

The spin density injection current $\dot{K}^{\mathrm{ab}}(\omega)$, where the
roman superscripts denote Cartesian components corresponding $\mathrm{a}$ to the
spin velocity direction and $\mathrm{b}$ to the spin polarization direction, can
be written in terms of a material response pseudotensor
$\mu^{\mathrm{abcd}}(\omega)$ as
\begin{equation*}
\dot{K}^{\mathrm{ab}}(\omega) = \mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega)
\label{eq:dotk}
\end{equation*}
where the pseudotensor is given by\cite{bhatPRL05}
\begin{widetext}
\begin{align}
\mu^{\mathrm{abcd}} (\omega) 
&=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}}
\sum_{vcc'}^{'}
\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'} 
\left( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
(c \leftrightarrow d)
\right) 
\right]
\delta(\omega-\omega_{cv})
\label{eq:mu}
\end{align}
\end{widetext}
where $K^{\mathrm{ab}}_{mn}(\mathbf{k}) =
\sum_{\ell}v^{\mathrm{a}}_{nl}(\mathbf{k}) S^{\mathrm{b}}_{lm}(\mathbf{k})$ are
the spin current matrix elements that, using time reversal invariance, satisfy
the relation $K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}(\mathbf{k})
$. The $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and
the sum only covers these states and since $\mu$ is real we have that
$\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$.

Now we define the spin velocity, $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ as
the speed at which the spin polarized in the $\mathrm{b}$   direction moves
along the $\mathrm{a}$ direction when a normal incident beam reaches the $xy$
plane with a polarization angle $\alpha$ as
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)},
\label{eq:vab}
\end{align}
\end{widetext}
where $\xi^{\mathrm{aa}}(\omega)$ are the carrier generation rate tensor
components\cite{arzatePRB14}.
% 
For an angle $\alpha = \frac{\pi}{4}$ this expression can be reduced to 
\begin{equation*}\label{eq:vab-90deg}
\mathcal{V}^{\mathrm{ab}} (\omega)
= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega) + \mu^{\mathrm{abyy}}(\omega) + 
2\mu^{\mathrm{abxy}}(\omega)}
{\xi^{\mathrm{xx}}(\omega) + \xi^{\mathrm{yy}}(\omega)}.
\end{equation*}

@@

Considering that we have 2D structures, one of the options to analyze the spin
velocity (Eq. \eqref{eq:vab}) is fixing the spin along the $x$, $y$, or $z$
Cartesian coordinates. We define the magnitude of the spin velocity with spin
polarization along the $\mathrm{b}$ direction as
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
&=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{align}
and the angle at which the spin velocity is directed on the $xy$ plane as
\begin{align}
\gamma_{\mathrm{b}} (\omega,\alpha)
&=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{align}
where the angle is measured in the counter-clockwise direction from the positive
$x$ Cartesian coordinate. We also define two special angles
\begin{align}
\gamma_{\mathrm{b \parallel}}(\omega,\alpha) &= \alpha, 
\label{eq:gamma-par} 
\\
\gamma_{\mathrm{b \perp}}(\omega,\alpha) &= \alpha \pm 90^{\circ}.
\label{eq:gamma-perp}
\end{align}
The first corresponds to the case when the spin velocity is directed, on the
$xy$ plane in the same direction of the polarization angle of the incoming beam;
the second one corresponds to the case when the spin velocity is directed
perpendicularly respect to the polarization angle of the incoming beam.

@@

Another possibility to analyze the spin velocity (Eq. \eqref{eq:vab}) is fixing
the velocity on the $xy$ plane along $x$ or $y$ Cartesian coordinate and then
define the magnitude of the spin velocity directed along the $\mathrm{a}$
direction as
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} .
},
\label{eq:vv-mag}
\end{align}
Then, the spin direction depends of the components of the previous equation and
so we define the spin orientation polar and azimuthal angles as
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}
where $\theta_{\mathrm{a}}(\omega,\alpha)$ is measured from the positive to the
negative $z$ Cartesian coordinate and $\varphi_{\mathrm{a}}(\omega,\alpha)$ is
measured on the $xy$ plane in the counter-clockwise direction from the positive
$x$ Cartesian coordinate.

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{alt} structure. Layer division, atom types and
positions for the \emph{alt} structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H & -0.61516 & -1.77416 &  0.73196 \\
1 & H &  0.61518 &  0.35514 &  0.73175 \\
2 & C & -0.61516 & -1.77264 & -0.49138 \\
2 & C & -0.61516 & -0.35600 & -0.72316 \\
2 & C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{up} structure. Layer division, atom types and
positions for the \emph{up} structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}

We preset the results for $|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ and
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ for the C$_{16}$H$_{8}$-alt
and C$_{16}$H$_{8}$-up structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. The \emph{alt}
system has alternating hydrogen atoms on the upper and bottom sides of the
carbon sheet, while the \emph{up} system has H only on the upper side. We take
the hexagonal carbon lattice to be on the $xy$ plane for both structures, and
the carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:alt-struc} and \ref{fig:up-struc}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA) with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used 
Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented in Eq.
\eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{alt} and \emph{up} structures are cutoff energies of 65\,Ha and
40\,Ha, respectively. The energy eigenvalues and matrix elements for the
\emph{up} ad \emph{alt} structures were calculated using 14452 $\mathbf{k}$
points and 8452 $\mathbf{k}$ points in the irreducible Brillouin zone (IBZ)
presenting LDA energy band gaps of 0.72\,eV and 0.088\,eV, respectively.
% 
We notice that within DFT, the LDA is only one of other possible methods that
can be used to determine the electronic structure of materials. Recent
investigations on graphene show some of the differences in calculated values
from several of these methods\cite{karamanisJPCC15,botelloACR14}. We note that
the LDA is as good as these other approaches. It is also known that the DFT
calculations predict a band gap for the material that differs from experiment.
This can be corrected using other \emph{ab initio} techniques, such as the GW
approximation\cite{onidaRMP02}, but this calculation has a very high
computational cost and is out of the scope in this paper. Even so, DFT still
remains as an effective and useful tool for computing diverse properties derived
from the electronic band structure.

@@

Using the Eq. \eqref{eq:vab}, we calculated the
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ response for the \emph{alt} and
\emph{up} 2D structures and for the CdSe and GaAs bulk systems and the results
are presented in Fig. \ref{fig:vab-str-comp}. The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.
% 
From the figure we have that the onset of the response starts when the energy of
the incoming beam is the same of the gap energy.
% 
The most intense response corresponds to the \emph{up} structure centered at
0.088\,eV corresponding to the Far Infrared (FIR) radiation and reaching a spin
velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the Near Infrared (NIR) and visible radiation, all the four
structures have contributions in the same order of magnitude.
% 
Starting with the 2D structures we have that the \emph{up} structure has other
two peaks centered at 1.94\,eV and 1.97\,eV reaching spin velocities of
22.2\,Km/s and -29.7\,Km/s, respectively, and the \emph{alt} structure has two
peaks centered at 0.72\,eV and 0.91\,eV reaching spin velocities of -40.2\,Km/s
and -32.9\,Km/s, respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting in
a spin velocity of -21.6\,Km/s.
% 
A negative quantity in the spin velocity means a change in the spin polarization
traveling in the opposite direction.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for the
2D and bulk structures. We found that the most intense response for the spin
velocity of the \emph{up} structure is 3.25 times more intense than CdSe and
4.03 times more intense than the GaAs bulk structures. Also, the \emph{alt}
structure has a response more intense than the bulk systems but being less
intense than the corresponding to the \emph{up} one.
%

@@

Using the Eq. \eqref{eq:vs-mag}, we calculated the
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ response and made the
analysis for the case when the spin is fixed in the $z$ direction, directed
perpendicularly to the surface of the \emph{up} and \emph{alt} structures. Also,
using the Eq. \eqref{eq:gamma-ang}, we determined the angle
$\gamma_{\mathrm{b}}(\omega,\alpha)$ where the spin-velocity is directed on the
surface of the each structure.
% 
For both cases, the most interesting case is that for which the spin is
perpendicular to the surface corresponding to $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$.

@@

We first analyzed two energy ranges for the \emph{up} structure, the first for
an incoming energy beam from 0.0\,eV to 0.2\,eV which include the THz and the
Mid Infrared (MIR) radiation, where the absolute maximum of the
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ response is obtained and
the second for an energy range from 1.80\,eV to 2.1\,eV, corresponding to
visible radiation, where two local maxima are found.
In Fig. \ref{fig:up-3d-vsz} we present the
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ spectra resulting from
evaluate Eq. \eqref{eq:vs-mag}.

@@

$45^{\circ}$. Also two local maxima are held for same polarization angles but
for an energy range of the incoming beam between 1.90\,eV and 2.05\,eV.

@@

fixing the energy of the incoming beam to 0.088\,eV and 0.912\,eV, respectively,
the corresponding velocity angle $\gamma_{\mathrm{z}}(\omega,\alpha)$ obtained
from evaluate Eq. \eqref{eq:gamma-ang}, and in the bottom frames the panels we
present the corresponding components $\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)$

>>>>

\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external
                        % documents and URLs 
\usepackage{blindtext}  % fill text
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
% \usepackage{showframe}
\usepackage[inline]{showlabels}

@@

\section{Introduction}

\blindtext
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/upstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/upstruc1}
    \caption{Up structure}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{figures/altstruc2}
    \\
    \includegraphics[width=\linewidth]{figures/altstruc1}
    \caption{Alt structure.}
    \label{fig:alt-struc}
\end{figure}
\blindtext
\blindtext

@@

The spin density injection current $\dot{K}^{\mathrm{ab}}(\omega)$, 
% 
{
\color{red}
where the
roman superscripts denote Cartesian components corresponding $\mathrm{a}$ to the
spin velocity direction and $\mathrm{b}$ to the spin polarization direction, can
be written in terms of a material response pseudotensor
$\mu^{\mathrm{abcd}}(\omega)$ as
}
\begin{equation*}
\dot{K}^{\mathrm{ab}}(\omega) = \mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega)
\label{eq:dotk}
\end{equation*}
where the pseudotensor is given by\cite{bhatPRL05}
\begin{widetext}
\begin{align}
\mu^{\mathrm{abcd}} (\omega) 
&=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}}
\sum_{vcc'}^{'}
\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'} 
\left( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
(c \leftrightarrow d)
\right) 
\right]
\delta(\omega-\omega_{cv})
\label{eq:mu}
\end{align}
\end{widetext}
{\color{red}
where $K^{\mathrm{ab}}_{mn}(\mathbf{k}) =
\sum_{\ell}v^{\mathrm{a}}_{nl}(\mathbf{k}) S^{\mathrm{b}}_{lm}(\mathbf{k})$ are
the spin current matrix elements that, using time reversal invariance, satisfy
the relation $K^{\mathrm{ab}}_{nm}(\mathbf{-k}) = K^{\mathrm{ab*}}(\mathbf{k})
$. 
}% 
The $'$ in the sum means that $c$ and $c'$ are quasi degenerate states and
the sum only covers these states and 
{\color{red}
since $\mu$ is real we have that
$\mu^{\mathrm{abcd}}(\omega) = \mu^{\mathrm{abdc}}(\omega)$.
}
Now we define the spin velocity, $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ as
the speed at which the spin polarized in the $\mathrm{b}$   direction moves
along the $\mathrm{a}$ direction when a normal incident beam reaches the $xy$
plane with a polarization angle $\alpha$ as
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)},
\label{eq:vab}
\end{align}
\end{widetext}
{\color{red}
where $\xi^{\mathrm{aa}}(\omega)$ are the carrier generation rate tensor
components\cite{arzatePRB14}.
}
% 
For an angle $\alpha = \frac{\pi}{4}$ this expression can be reduced to 
\begin{equation*}\label{eq:vab-90deg}
\mathcal{V}^{\mathrm{ab}} (\omega)
= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega) + \mu^{\mathrm{abyy}}(\omega) + 
2\mu^{\mathrm{abxy}}(\omega)}
{\xi^{\mathrm{xx}}(\omega) + \xi^{\mathrm{yy}}(\omega)}.
\end{equation*}

@@

{\color{red} Considering that we have 2D structures, one of the options to
analyze the spin velocity (Eq. \eqref{eq:vab}) is fixing the spin parallel to
the surface of the structure in the $x$ or $y$ directions or perpendicularly to
it in the $z$ direction. We define the magnitude of the spin velocity with spin
polarization along the $\mathrm{b}$ Cartesian coordinate as}
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
&=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{align}
and the angle at which the spin velocity is directed on the $xy$ plane as
\begin{align}
\gamma_{\mathrm{b}} (\omega,\alpha)
&=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{align}
where the angle is measured in the counter-clockwise direction from the positive
$x$ Cartesian coordinate. 
{\color{red}
We also define two special angles
\begin{align}
\gamma_{\mathrm{b \parallel}}(\omega,\alpha) &= \alpha, 
\label{eq:gamma-par} 
\\
\gamma_{\mathrm{b \perp}}(\omega,\alpha) &= \alpha \pm 90^{\circ}.
\label{eq:gamma-perp}
\end{align}
The first corresponds to the case when the spin velocity is directed, on the
$xy$ plane in the same direction of the polarization angle of the incoming beam;
the second one corresponds to the case when the spin velocity is directed
perpendicularly respect to the polarization angle of the incoming beam.
}

@@

{\color{red} Another possibility to analyze the Eq. \eqref{eq:vab} is fixing the velocity on
the $xy$ plane along $x$ or $y$ Cartesian coordinate and then define the
magnitude of the spin velocity directed along the $\mathrm{a}$ direction as}
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} .
},
\label{eq:vv-mag}
\end{align}
{\color{red} Then, the spin direction depends of the $x$, $y$, and $z$
components of the previous equation and so we define the spin orientation polar
and azimuthal angles as}
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}
{\color{red} where $\theta_{\mathrm{a}}(\omega,\alpha)$ is measured from the
positive to the negative $z$ Cartesian coordinate and
$\varphi_{\mathrm{a}}(\omega,\alpha)$ is measured on the $xy$ plane in the
counter-clockwise direction from the positive $x$ Cartesian coordinate.}

@@

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H & -0.61516 & -1.77416 &  0.73196 \\
1 & H &  0.61518 &  0.35514 &  0.73175 \\
2 & C & -0.61516 & -1.77264 & -0.49138 \\
2 & C & -0.61516 & -0.35600 & -0.72316 \\
2 & C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{up} structure. Layer division, atom types and
positions for the \emph{up} structure. The structure unit cell was divided in
two layers corresponding to hydrogen and carbon atoms.The corresponding layer
atom position is depicted in Fig. \ref{fig:up-struc} with the corresponding
number of layer.}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Layer \quad & \quad Atom \qquad & \multicolumn{3}{c}{Position [\AA]} \\
\cline{3-5}
\quad No.   \quad & \quad type \qquad & $x$ & $y$ & $z$  \\
\hline
1 & H &  -0.61516 &  -1.42140 & \ 1.47237 \\
2 & C &  -0.61516 &  -1.73300 & \ 0.39631 \\
3 & C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
4 & C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
5 & C &  -0.61516 &  -0.37396 &  -0.39632 \\
6 & H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}

\caption{Unit cell of \emph{alt} structure. Layer division, atom types and
positions for the \emph{alt} structure. The structure unit cell was divided in
six layers corresponding each one to atoms in different $z$ positions. The
corresponding layer atom position is depicted in Fig. \ref{fig:alt-struc} with
the corresponding number of layer.}
\label{tab:alt-unitcell}
\end{table}

We preset the results for $|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ and
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ for the C$_{16}$H$_{8}$-alt
and C$_{16}$H$_{8}$-up structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. The \emph{up}
structure has hydrogen atoms only on the upper side of the carbon sheet while
the \emph{alt} structure has alternating hydrogen atoms on the upper and bottom
sides. We take the hexagonal carbon lattice to be on the $xy$ plane for both
structures, and the carbon-hydrogen bonds on the perpendicular $xz$ plane, as
depicted in Figs. \ref{fig:alt-struc} and \ref{fig:up-struc}
% 
{\color{blue} and the coordinates for the \emph{up} and \emph{alt} unit cells of
the structures are presented in Tables \ref{tab:up-unitcell} and
% 
\ref{tab:alt-unitcell}. In same tables we present the layer division needed to
calculate the layer-by-layer contribution for the
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ and
$|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ presented in Eqns. 
% 
\eqref{eq:vs-layer} and \eqref{eq:vv-layer}. The \emph{up} structure was divided
in two layers, the first comprised of the top hydrogen atoms denoted by the
number 1 in Table \ref{tab:up-unitcell} and in the Fig. \ref{fig:up-struc} and
the second comprised of carbon atoms and denoted by the number 2. The \emph{alt}
structure was divided in six layers denoted with numbers from 1 to 6 in Table
\ref{tab:alt-unitcell} and in Fig. \ref{fig:alt-struc}. The first and sixth
layers correspond to hydrogen atoms in the top and bottom positions and from the
second to the fifth correspond to carbon atoms placed in different positions.}

{\color{red} We calculated the self-consistent ground state and the Kohn-Sham
states using density functional theory in the local density approximation (DFT-
LDA) with a planewave basis using the ABINIT code \cite{gonzeCPC09}.}
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq.
\eqref{eq:mu}.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{alt} and \emph{up} structures are cutoff energies of 65\,Ha and
40\,Ha, respectively. The energy eigenvalues and matrix elements for the
\emph{up} ad \emph{alt} structures were calculated using 14452 $\mathbf{k}$
points and 8452 $\mathbf{k}$ points in the irreducible Brillouin zone (IBZ)
{\color{red} resulting in LDA energy band gaps of 0.72\,eV and 0.088\,eV,
respectively.
% 
We notice that within DFT, the LDA is only one of other possible methods that
can be used to determine the electronic structure of materials. Recent
investigations on graphene show some of the differences in calculated values
from several of these methods\cite{karamanisJPCC15,botelloACR14}. We note that
the LDA is as good as these other approaches. It is also known that the DFT
calculations predict a band gap for the material that differs from experiment.
This can be corrected using other \emph{ab initio} techniques, such as the GW
approximation\cite{onidaRMP02}, but this calculation has a very high
computational cost and is out of the scope in this paper. Even so, DFT still
remains as an effective and useful tool for computing diverse properties derived
from the electronic band structure.}

@@

{\color{red}Using the Eq. \eqref{eq:vab}, we calculated the
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ response for the \emph{up} and
\emph{alt} 2D structures and for the CdSe and GaAs bulk systems; the results
are presented in Fig. \ref{fig:vab-str-comp}. The angle $\alpha$ presented in
the response of each structure is that for which the response is maximized in
each case.}
% 
From the figure we have that the onset of the response starts when the energy of
the incoming beam is the same of the gap energy.
% 
The most intense response corresponds to the \emph{up} structure centered at
0.088\,eV corresponding to the {\color{red} Mid Infrared (MIR)} radiation and
reaching a spin velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the {\color{red} Near Infrared (NIR) to visible radiation}, all the four
structures have contributions in the same order of magnitude.
% 
For this energy range the \emph{up} structure has two peaks centered at 1.94\,eV
and 1.97\,eV reaching spin velocities of 22.2\,Km/s and -29.7\,Km/s,
respectively, and the \emph{alt} structure has two peaks centered at 0.72\,eV
and 0.91\,eV reaching spin velocities of -40.2\,Km/s and -32.9\,Km/s,
respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting in
a spin velocity of -21.6\,Km/s.
% 
{\color{red} A negative quantity in the spin velocity means a change in the spin
polarization traveling in the opposite direction.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for the
2D and bulk structures. We found that the most intense response for the spin
velocity corresponds to the \emph{up} structure being 3.25 times more intense
than that of the CdSe and 4.03 times more intense than that of the GaAs bulk
structures. Also, the \emph{alt} structure has a response more intense than the
bulk systems but being less intense than the corresponding to the \emph{up}
one.}
%

@@

{\color{red} Using the Eq. \eqref{eq:vs-mag}, we calculated the
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ response and made the
analysis for the case when the spin is fixed in the $z$ coordinate, directed
perpendicularly to the surface of the \emph{up} and \emph{alt} structures. Also,
using the Eq. \eqref{eq:gamma-ang}, we determined the angle
$\gamma_{\mathrm{b}}(\omega,\alpha)$ where the spin-velocity is directed on the
surface of the each structure.}
%

@@

We first analyzed two energy ranges in Fig. \ref{fig:up-3d-vsz} for the
\emph{up} structure, the first for an incoming energy beam from 0.0\,eV to
0.2\,eV (top panel)  which include the THz and the Mid Infrared (MIR) radiation,
where the absolute maximum of the
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ response is obtained, and
the second for an energy range from 1.80\,eV to 2.1\,eV (bottom panel),
corresponding to visible radiation, where two local maxima are found.

@@

$45^{\circ}$. Also the two local maxima are held for same beam polarization
angles but for an energy range between 1.90\,eV and 2.05\,eV.
fixing the energy of the incoming beam to 0.088\,eV and 1.972\,eV, respectively,
the corresponding velocity angle $\gamma_{\mathrm{z}}(\omega,\alpha)$, and in
the bottom frames of the panels we present the corresponding components
$\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)$ and
$\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)$. Also we present two circled and

@@
